An Evolutionary Algorithm for the Multiobjective Risk-Equity Constrained Routing Problem

### 2011-11-29 DAW_hazmatHypE/trunk/paper/hazmatRouting_EvolOptim.tex
new figure for comparing W-Hype with HypE and corresponding (preliminary) text

\subsection{Implementation in PISA}
The proposed algorithm has been implemented within the multiobjective optimization framework PISA \citep{bltz2003a} and is going to be made available via the official PISA web page\footnote{Until the source code is available via the PISA web page \url{http://www.tik.ee.ethz.ch/pisa/}, we made it temporarily available at \url{http://researchers.lille.inria.fr/~brockhof/downloads/hazmatvariator20111110.zip} as supplementary material for the review process.}. More precisely, the implementation of an algorithm/problem combination in PISA is divided into the problem-specific part (called \emph{variator module}) consisting of representation, initialization, and variation operators and the problem-independent part (called the \emph{selector module}). As selection module, any of the state-of-the-art selectors in PISA can be used off-the-shelf. In this study, we choose to use HypE \citep{bz2011a}, NSGA-II \citep{}, and W-HypE \citep{} with a focus on the first one.

@@

\begin{table}
\centering
\caption{\label{tab:referencepoints} Used reference points within the environmental selection of HypE}
\vspace{1em}
\begin{tabular}{lr}
	\toprule
	instance & reference point\\ \midrule
	\# commodities: 2 & $(10^8, 10^8, 10^8)$\\
	\# commodities: 3 & $(10^8, 10^8, 10^8)$\\
	\# commodities: 4 & $(10^9, 10^9, 10^9)$\\
	\bottomrule
\end{tabular}
\end{table}

@@

\begin{figure}%
	\centering
	\includegraphics[width=0.48\columnwidth]{../experiments/randVsCost/hypervolumes/hypervolumeOverTime_50.eps}%
	\includegraphics[width=0.48\columnwidth]{../experiments/randVsCost/hypervolumes/hypervolumeOverTime_100.eps}%
	\caption{\label{fig:hypervolumeOverTime} Boxplots of hypervolume values over time for numbers of generations in $[250, \ldots, 2000]$ for HypE with population size 50 (left) and population size 100 (right) on the lazio\_4 instance.}
\end{figure}

@@

\begin{figure}%
	\includegraphics[width=\columnwidth]{../experiments/randVsCost/allsolutions.eps}%
	\caption{\label{fig:allsolutions} Comparison between the random initialization (circles) and the cost-optimal one (crosses). Shown are the populations of 10 independent HypE runs after 1000 generations for all three objectives (top left) and each combination of two objectives.}
\end{figure}

\begin{figure}%
	\includegraphics[width=\columnwidth]{../experiments/randVsCost/onlynondominated.eps}%
	\caption{\label{fig:onlynondominated} Comparison between the random initialization (circles) and the cost-optimal one (crosses). Shown are only the non-dominated solutions in the final populations of 10 independent HypE runs after 1000 generations.}
\end{figure}

@@

\subsection{Changing the Environmental Selection}
With a last experiment, we show the influence of the algorithm's environmental selection scheme on the performance and in particular how this can be used to steer the search towards specific regions of the Pareto front.
\TODO{write}

>>>>

\subsection{Implementation in PISA}
The proposed algorithm has been implemented within the multiobjective optimization framework PISA \citep{bltz2003a} and is going to be made available via the official PISA web page\footnote{Until the source code is available via the PISA web page \url{http://www.tik.ee.ethz.ch/pisa/}, we made it temporarily available at \url{http://researchers.lille.inria.fr/~brockhof/downloads/hazmatvariator20111110.zip} as supplementary material for the review process.}. More precisely, the implementation of an algorithm/problem combination in PISA is divided into the problem-specific part (called \emph{variator module}) consisting of representation, initialization, and variation operators and the problem-independent part (called the \emph{selector module}). As selection module, any of the state-of-the-art selectors in PISA can be used off-the-shelf. In this study, we choose to use HypE \citep{bz2011a}, NSGA-II \citep{dapm2002a}, and W-HypE \citep{} with a focus on the first one.

@@

@@

\begin{figure}
	\centering
	\includegraphics[width=0.48\columnwidth]{../experiments/randVsCost/hypervolumes/hypervolumeOverTime_50.eps}%
	\includegraphics[width=0.48\columnwidth]{../experiments/randVsCost/hypervolumes/hypervolumeOverTime_100.eps}%
	\caption{\label{fig:hypervolumeOverTime} Boxplots of hypervolume values over time for numbers of generations in $[250, \ldots, 2000]$ for HypE with population size 50 (left) and population size 100 (right) on the lazio\_4 instance.}
\end{figure}

@@

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{../experiments/randVsCost/allsolutions.eps}%
	\caption{\label{fig:allsolutions} Comparison between the random initialization (circles) and the cost-optimal one (crosses). Shown are the populations of 10 independent HypE runs after 1000 generations for all three objectives (top left) and each combination of two objectives.}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{../experiments/randVsCost/onlynondominated.eps}%
	\caption{\label{fig:onlynondominated} Comparison between the random initialization (circles) and the cost-optimal one (crosses). Shown are only the non-dominated solutions in the final populations of 10 independent HypE runs after 1000 generations.}
\end{figure}

@@

\subsection{Changing the Environmental Selection}
With a last experiment, we show the influence of the algorithm's environmental selection scheme on the performance and in particular how this can be used to steer the search towards specific regions of the Pareto front.

Instead of the environmental selection of HypE \citep{bz2011a}, other selection schemes can be easily used with the PISA framework and we show the resutls of two other algorithms exemplary. On the one hand, we employ the famous NSGA-II selection scheme \citep{dapm2002a} and, on the other hand, we use a generalized version of HypE towards a generalized weighted hypervolume indicator \citep{abbz2009b,bbtz2011a}. The boxplots of the resulting hypervolume indicator values after 100,000 function evaluations (population of size 100 after 1000 generations) are shown in Fig.~\ref{fig:algoComparison}. \TODO{What can be seen is that both? algorithms show a reduced performance regarding the standard hypervolume value. For NSGA-II, this is expected as its environmental selection schemes becomes less effective than in bi-objective problems when the number of objectives increases \citep{wbn2007a,bz2011a}. This is not the case for indicator-based algorithms such as HypE which were specifically proposed to circumvent the problems of algorithms optimizing non-compliant quality measures such as NSGA-II.

The reason for the reduced performance in the standard hypervolume indicator for W-Hype, however, is a different one. W-HypE was developed to steer the search towards regions of the objective space that have been specified by the user. To this end, it does not optimize the standard hypervolume indicator but a more generalized version of it, the weighted hypervolume indicator \citep{zbt2007a}. The idea behind W-HypE is that solutions in objective space regions with a higher weight will get a higher fitness due to the \emph{weighted} hypervolume they dominate. In order to be efficient also for a large number of objectives, W-HypE does not calculate the weighted hypervolume indicator values exactly (which is not possible for most of the weight functions) but instead samples them. Here, 10,000 samples are used in each iteration of the algorithm.

To show the principles of W-HypE in the application of our hazmat routing problem, we defined the weight of W-HypE's underlying weighted hypervolume indicator as constant in a box with corners $[7\cdot 10^3, 3\cdot 10^7, 5\cdot 10^5]$ and $[10^4, 4\cdot 10^7, 1.1\cdot 10^6]$ and zero otherwise. Instead of optimizing the standard hypervolume indicator as HypE, W-HypE now optimizes this weighted hypervolume indicator, resulting in better, i.e., dominating solutions in regions in or close to the box. Figure~\ref{fig:whype} shows the resulting non-dominated solutions of both algorithms with population size 25 after 10 independent runs of 4,000 generations each as well as the sampling box.}

This advantage of W-HypE over HypE within the pre-defined box can be observed both in early and late generations, cp.\ Fig.~\ref{fig:whype}.

\TODO{write}

\begin{figure}
	\centering
	\includegraphics[width=0.8\columnwidth]{../experiments/randVsCost/WHypEVsHypE_ns4_1_OriginalCosts2.eps}%
	\caption{\label{fig:whype} Comparison between the standard HypE selection ($\circ$) with the one of W-HypE ($\times$) when the weighted hypervolume indicator is only sampled in the gray shaded box $[7\cdot 10^3, 10^4]\times[3\cdot 10^7, 4\cdot 10^7]\times[5\cdot 10^5, 1.1\cdot 10^6]$. Shown are all non-dominated solutions for each algorithm after 10 runs with 4,000 generations and population size 25.}
\end{figure}
